Identifying Factors That Affect Political Freedom Within the OECD

### 2020-11-29 paper/paper.Rmd
referenced index
.

```{r setup}
library(tidyverse)
library(kableExtra)
library(stargazer)
library(patchwork)
size <- 30
library("papaja")
r_refs("r-references.bib")
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message = FALSE, 
                      fig.width = 12, fig.height = 8) 
```

```{r, cache=TRUE}
# Read and Clean ----------------------------------------------------------
files <- list.files(path = here::here("raw_data"), pattern = ".csv")
cleaning <- function(df){
  df <- pivot_longer(df, cols = -country, names_to = "year")
}
data <- files %>% 
  map(function(x) read_csv(paste0("raw_data/", x))) %>% 
  setNames(gsub("\\.csv$", "", files)) %>% 
  map(cleaning) %>% 
  bind_rows(.id = "id") %>% 
  pivot_wider(names_from = id)

# Filtering ---------------------------------------------------------------
countries <- readRDS(here::here("data", "countries.RDS"))
data <- data %>%
  mutate(year = as.numeric(year)) %>% 
  filter(year >= 2000, year < 2019, country %in% countries) %>% 
  group_by(country) %>%
  mutate_at(vars(-country),list(~ifelse(is.na(.), median(., na.rm = TRUE), .))) %>% 
# Iceland spent 0 so I needed to manually recode that
  mutate(military_spending_pct_of_gdp = replace_na(military_spending_pct_of_gdp, 0),
  murder_per_mil_people = replace(murder_per_mil_people, country == "Mexico", 29.07),
         murder_per_mil_people = replace(murder_per_mil_people, country == "Chile", 4.4),
         murder_per_mil_people = replace(murder_per_mil_people, country == "Colombia", 25.34)) %>% 
  relocate(polrights_fh) %>% 
  mutate(polrights_fh = (8 - polrights_fh),
 military_spending_pct_of_gdp = military_spending_pct_of_gdp * 100) %>% 
  mutate(corruption_perception_index_cpi = (100 - corruption_perception_index_cpi)) %>% 
  ungroup()
# Best Subset Selection ---------------------------------------------------------------
vars <- data %>% 
  select(-country, -year, -ends_with("_fh")) %>%
  names()
models <- list()
for (i in 1:length(vars)) {
  vc <- combn(vars,i)
  for (j in 1:ncol(vc)) {
    model <- as.formula(paste0("polrights_fh", " ~", paste0(vc[,j], collapse = " + ")))
    models <- c(models, model)
    }
  }
subsets <- map(models, function(x) lm(x, data)) %>% 
  map(broom::glance) %>% 
  setNames(models) %>% 
  bind_rows(.id = "id") %>% 
  rename(model = id) %>% 
  mutate(Model = str_replace_all(model, "_", " "),
         Model = str_replace(Model, "~", "="), 
         Model = str_to_title(Model))
```

@@

2020 is an election year, which has many of us reflecting on political freedom. Political freedoms are important because they have a dramatic impact on the quality of life for the citizens of a country. Such reflection has us asking: What factors are associated with an increase in political freedom around the world? Political freedom is an abstract concept that does not lend well to quantification. However, Freedomhouse [@Freedomhouse], a U.S. based think tank and research institute, attempts this by assigning an index of political rights to each country every year based on a fixed criteria. Using the Papaja Package by [@R-papaja] we take Rmarkdown directly to APA format.

@@

These data sets were separate so we downloaded them individually, cleaned and combined them together using R [@R-base] and the Tidyverse Package [@R-tidyverse]. This resulted in a data set with all 196 countries from years 1950 to 2030 (data after 2019 were projections). It was immediately apparent that there was a substantial amount of missing data in an obvious pattern. Less developed countries tended to have more missing data points, and the further back in time the more missing data. Had we conducted this study as it was it would have been severely biased towards more developed countries as missing data points would be dropped. This forced us to narrow the scope of the study and focus only on countries within the Organization for Economic Co-operation and Development (OECD). This subset of countries has much more complete data. Therefore, we removed data from countries that are not members of the OECD and selected only the years 2000 through 2018.

>>>>

```{r setup}
library(tidyverse)
library(kableExtra)
library(stargazer)
library(patchwork)
library("papaja")
size <- 30
r_refs("r-references.bib")
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message = FALSE, 
                      fig.width = 12, fig.height = 8) 
```

```{r, cache=TRUE}
# Read and Clean ----------------------------------------------------------
files <- list.files(path = here::here("raw_data"), pattern = ".csv")
cleaning <- function(df){
  df <- pivot_longer(df, cols = -country, names_to = "year")
}
data <- files %>% 
  map(function(x) read_csv(paste0("raw_data/", x))) %>% 
  setNames(gsub("\\.csv$", "", files)) %>% 
  map(cleaning) %>% 
  bind_rows(.id = "id") %>% 
  pivot_wider(names_from = id)
# Filtering ---------------------------------------------------------------
countries <- readRDS(here::here("data", "countries.RDS"))
data <- data %>%
  mutate(year = as.numeric(year)) %>% 
  filter(year >= 2000, year < 2019, country %in% countries) %>% 
  group_by(country) %>%
  mutate_at(vars(-country),list(~ifelse(is.na(.), median(., na.rm = TRUE), .))) %>% 
# Iceland spent 0 so I needed to manually recode that
  mutate(military_spending_pct_of_gdp = replace_na(military_spending_pct_of_gdp, 0),
  murder_per_mil_people = replace(murder_per_mil_people, country == "Mexico", 29.07),
         murder_per_mil_people = replace(murder_per_mil_people, country == "Chile", 4.4),
         murder_per_mil_people = replace(murder_per_mil_people, country == "Colombia", 25.34)) %>% 
  relocate(polrights_fh) %>% 
  mutate(polrights_fh = (8 - polrights_fh),
 military_spending_pct_of_gdp = military_spending_pct_of_gdp * 100) %>% 
  mutate(corruption_perception_index_cpi = (100 - corruption_perception_index_cpi)) %>% 
  ungroup()
# Best Subset Selection ---------------------------------------------------------------
vars <- data %>% 
  select(-country, -year, -ends_with("_fh")) %>%
  names()
models <- list()
for (i in 1:length(vars)) {
  vc <- combn(vars,i)
  for (j in 1:ncol(vc)) {
    model <- as.formula(paste0("polrights_fh", " ~", paste0(vc[,j], collapse = " + ")))
    models <- c(models, model)
    }
  }
subsets <- map(models, function(x) lm(x, data)) %>% 
  map(broom::glance) %>% 
  setNames(models) %>% 
  bind_rows(.id = "id") %>% 
  rename(model = id) %>% 
  mutate(Model = str_replace_all(model, "_", " "),
         Model = str_replace(Model, "~", "="), 
         Model = str_to_title(Model))
```

@@

2020 is an election year, which has many of us reflecting on political freedom. Political freedoms are important because they have a dramatic impact on the quality of life for the citizens of a country. Such reflection has us asking: What factors are associated with an increase in political freedom around the world? Political freedom is an abstract concept that does not lend well to quantification. However, Freedomhouse [@Freedomhouse], a U.S. based think tank and research institute, attempts this by assigning an index of political rights to each country every year based on a fixed criteria. Using the Papaja Package [@R-papaja] we take Rmarkdown directly to APA format.

@@

These data sets were separate so we downloaded them individually, cleaned and combined them together using R [@R-base] and the Tidyverse Package [@R-tidyverse]. We included all the code in the appendix. This resulted in a data set with all 196 countries from years 1950 to 2030 (data after 2019 were projections). It was immediately apparent that there was a substantial amount of missing data in an obvious pattern. Less developed countries tended to have more missing data points, and the further back in time the more missing data. Had we conducted this study as it was it would have been severely biased towards more developed countries as missing data points would be dropped. This forced us to narrow the scope of the study and focus only on countries within the Organization for Economic Co-operation and Development (OECD). This subset of countries has much more complete data. Therefore, we removed data from countries that are not members of the OECD and selected only the years 2000 through 2018.
